Time invariant analysis of epidemics with \pkg{EpiCompare}

### 2021-05-22 jss/EpiCompare-paper/EpiCompare-paper.Rmd
second round of minor edits to 3's final subpart

\ben{\sout{Finally,} Often}\footnote{\ben{[Ben says: I feel like starting with "finally" causes a weird pause - but maybe that's just me.]}} the data analysis pipeline \ben{\sout{often}} ends with plots, tables, and summary statistics \ben{\sout{that} which} are used to \sout{\ben{compare} \shannon{and assess}} \ben{\sout{different models, and simulations to one another} assess model preformance and compare across models or simulations.} In \pkg{EpiCompare} we provide a set of comparison and assessment tools for model and simulation results that extend beyond the standard performance metrics (e.g. mean squared error or AIC) and into the lens of time-invariant analysis.  We have found \ben{that} these tools \ben{\sout{to be} are} specifically applicable for situations where only one season or cycle of an epidemic has occurred, \ben{or is the object of interest}.

\ben{The first set of of tools surround the creation of prediction regions. We can create a prediction regions from model simulations to examine if our model simulations capture the true epidemic trajectory. We do so in a time-invariant way and utlize a filamental representation of the model simulations and the true epidemic.} [For three-state epidemic models, we provide the `ggplot`/`ggtern` extension `geom_prediction_band()` which creates a prediction region around the top $1-\alpha$ proportion of the simulations. In this visual setting, comparing this prediction to the true epidemic trajectory or comparing the prediction regions defined by two different models' simulations can be done by eye. In \pkg{EpiCompare}, we also provide these prediction regions for epidemic models with with more than three states. The functions `create_convex_hull_structure()` and `create_delta_ball_structure()` create different geometric representations of prediction regions for any dimensional state-based model. For both of these geometric structures, we provide functions to check if a path is contained (`contained()`)]. \ben{We can also use these prediction regions to visually or mathematically compare how similar two sets of simulations are. In \pkg{EpiCompare} we provide the} `hausdorff_dist()` \ben{function to calculate the Hausdorff distance between multiple prediction regions.}

[We also provide functions to calculate the "extremeness" of a true epidemic trajectory compared to simulated epidemics via the equi-distance filamental trajectory representation as mentioned in Section \ref{sec:beyond-r0-sir}.] \ben{We provide implementations of a few score functions that capture extremeness of epidemics relative to other epidemics, and these scores can be turned into an extremeness measure with} `mean(sim_scores > truth_score)`. [Specifically, functions like  `distance_pseudo_density_function()` can calculate a pseudo-density estimate of the true epidemic relative to simulated ones. Functions `distance_depth_function()` and `local_distance_depth_function()` provide depth scores that suggest how geometrically central an epidemic is to simulations.]

>>>>

\ben{\sout{Finally,} Often}\footnote{\ben{[Ben says: I feel like starting with "finally" causes a weird pause - but maybe that's just me.]}} the data analysis pipeline \ben{\sout{often}} ends with plots, tables, and summary statistics that are used to \sout{\ben{compare} \shannon{and assess}} \ben{\sout{different models, and simulations to one another} assess model preformance and compare across models or simulations.} In \pkg{EpiCompare} we provide a set of comparison and assessment tools for model and simulation results that extend beyond the standard performance metrics (e.g. mean squared error or AIC) and into the lens of time-invariant analysis.  We have found \ben{that} these tools \ben{\sout{to be} are} specifically applicable for situations where only one season or cycle of an epidemic has occurred, \ben{or is the object of interest}.

\footnote{\ben{[Ben says: This changed introduction was done to highlight that we can use prediction regions for mutiple types of comparisons.]}}\ben{The first set of of tools surround the creation of prediction regions. We can create a prediction regions from model simulations to examine if our model simulations capture the true epidemic trajectory. We do so in a time-invariant way and utilizing filamental representations of the model simulations and the true epidemic.} [For three-state epidemic models, we provide the `ggplot`/`ggtern` extension `geom_prediction_band()` which creates a prediction region around the top $1-\alpha$ proportion of the simulations. In this visual setting, comparing this prediction \ben{region} to the true epidemic trajectory \ben{\sout{or comparing the prediction regions defined by two different models' simulations}}\footnote{\ben{[Ben says: this part will be highlighted in later sentences.]}} can be done by eye. In \pkg{EpiCompare}, we also provide these prediction regions for epidemic models with with more than three states. The functions `create_convex_hull_structure()` and `create_delta_ball_structure()` create different geometric representations of prediction regions for any dimensional state-based model. For both of these geometric structures, we provide functions to check if a path is contained (`contained()`)]. \ben{We can also use these prediction regions to visually or mathematically compare how similar two sets of simulations are. In \pkg{EpiCompare} we provide the} `hausdorff_dist()` \ben{function to calculate the Hausdorff distance between multiple prediction regions, when visual comparison is not possible.}

[We also provide functions to calculate the "extremeness" of a true epidemic trajectory compared to simulated epidemics via the equi-distance filamental trajectory representation as mentioned in Section \ref{sec:beyond-r0-sir}.] \ben{We provide implementations of a few distance-based score functions that capture how "reasonable" an epidemic is relative to other epidemics, and these scores can be turned into an extremeness measure with} `mean(sim_scores > truth_score)`. [Specifically, functions like  `distance_pseudo_density_function()` can calculate a pseudo-density estimate of the true epidemic relative to simulated ones. Functions `distance_depth_function()` and `local_distance_depth_function()` provide depth scores that suggest how geometrically central an epidemic is to simulations.]

### 2021-05-22 jss/EpiCompare-paper/EpiCompare-paper.tex
second round of minor edits to 3's final subpart

\textcolor{violet}{\sout{Finally,} Often}\footnote{\textcolor{violet}{[Ben says: I feel like starting with "finally" causes a weird pause - but maybe that's just me.]}}
the data analysis pipeline \textcolor{violet}{\sout{often}} ends with
plots, tables, and summary statistics
\textcolor{violet}{\sout{that} which} are used to
\sout{\textcolor{violet}{compare} \textcolor{orange}{and assess}}
\textcolor{violet}{\sout{different models, and simulations to one another} assess model preformance and compare across models or simulations.}
In \pkg{EpiCompare} we provide a set of comparison and assessment tools
for model and simulation results that extend beyond the standard
performance metrics (e.g.~mean squared error or AIC) and into the lens
of time-invariant analysis. We have found \textcolor{violet}{that} these
tools \textcolor{violet}{\sout{to be} are} specifically applicable for
situations where only one season or cycle of an epidemic has occurred,
\textcolor{violet}{or is the object of interest}.

\textcolor{violet}{The first set of of tools surround the creation of prediction regions. We can create a prediction regions from model simulations to examine if our model simulations capture the true epidemic trajectory. We do so in a time-invariant way and utlize a filamental representation of the model simulations and the true epidemic.}
{[}For three-state epidemic models, we provide the
\texttt{ggplot}/\texttt{ggtern} extension
\texttt{geom\_prediction\_band()} which creates a prediction region
around the top \(1-\alpha\) proportion of the simulations. In this
visual setting, comparing this prediction to the true epidemic
trajectory or comparing the prediction regions defined by two different
models' simulations can be done by eye. In \pkg{EpiCompare}, we also
provide these prediction regions for epidemic models with with more than
three states. The functions \texttt{create\_convex\_hull\_structure()}
and \texttt{create\_delta\_ball\_structure()} create different geometric
representations of prediction regions for any dimensional state-based
model. For both of these geometric structures, we provide functions to
check if a path is contained (\texttt{contained()}){]}.
\textcolor{violet}{We can also use these prediction regions to visually or mathematically compare how similar two sets of simulations are. In \pkg{EpiCompare} we provide the}
\texttt{hausdorff\_dist()}
\textcolor{violet}{function to calculate the Hausdorff distance between multiple prediction regions.}

{[}We also provide functions to calculate the ``extremeness'' of a true
epidemic trajectory compared to simulated epidemics via the
equi-distance filamental trajectory representation as mentioned in
Section \ref{sec:beyond-r0-sir}.{]}
\textcolor{violet}{We provide implementations of a few score functions that capture extremeness of epidemics relative to other epidemics, and these scores can be turned into an extremeness measure with}
\texttt{mean(sim\_scores\ \textgreater{}\ truth\_score)}.
{[}Specifically, functions like
\texttt{distance\_pseudo\_density\_function()} can calculate a
pseudo-density estimate of the true epidemic relative to simulated ones.
Functions \texttt{distance\_depth\_function()} and
\texttt{local\_distance\_depth\_function()} provide depth scores that
suggest how geometrically central an epidemic is to simulations.{]}

>>>>

\textcolor{violet}{\sout{Finally,} Often}\footnote{\textcolor{violet}{[Ben says: I feel like starting with "finally" causes a weird pause - but maybe that's just me.]}}
the data analysis pipeline \textcolor{violet}{\sout{often}} ends with
plots, tables, and summary statistics that are used to
\sout{\textcolor{violet}{compare} \textcolor{orange}{and assess}}
\textcolor{violet}{\sout{different models, and simulations to one another} assess model preformance and compare across models or simulations.}
In \pkg{EpiCompare} we provide a set of comparison and assessment tools
for model and simulation results that extend beyond the standard
performance metrics (e.g.~mean squared error or AIC) and into the lens
of time-invariant analysis. We have found \textcolor{violet}{that} these
tools \textcolor{violet}{\sout{to be} are} specifically applicable for
situations where only one season or cycle of an epidemic has occurred,
\textcolor{violet}{or is the object of interest}.

\footnote{\textcolor{violet}{[Ben says: This changed introduction was done to highlight that we can use prediction regions for mutiple types of comparisons.]}}\textcolor{violet}{The first set of of tools surround the creation of prediction regions. We can create a prediction regions from model simulations to examine if our model simulations capture the true epidemic trajectory. We do so in a time-invariant way and utilizing filamental representations of the model simulations and the true epidemic.}
{[}For three-state epidemic models, we provide the
\texttt{ggplot}/\texttt{ggtern} extension
\texttt{geom\_prediction\_band()} which creates a prediction region
around the top \(1-\alpha\) proportion of the simulations. In this
visual setting, comparing this prediction \textcolor{violet}{region} to
the true epidemic trajectory
\textcolor{violet}{\sout{or comparing the prediction regions defined by two different models' simulations}}\footnote{\textcolor{violet}{[Ben says: this part will be highlighted in later sentences.]}}
can be done by eye. In \pkg{EpiCompare}, we also provide these
prediction regions for epidemic models with with more than three states.
The functions \texttt{create\_convex\_hull\_structure()} and
\texttt{create\_delta\_ball\_structure()} create different geometric
representations of prediction regions for any dimensional state-based
model. For both of these geometric structures, we provide functions to
check if a path is contained (\texttt{contained()}){]}.
\textcolor{violet}{We can also use these prediction regions to visually or mathematically compare how similar two sets of simulations are. In \pkg{EpiCompare} we provide the}
\texttt{hausdorff\_dist()}
\textcolor{violet}{function to calculate the Hausdorff distance between multiple prediction regions, when visual comparison is not possible.}

{[}We also provide functions to calculate the ``extremeness'' of a true
epidemic trajectory compared to simulated epidemics via the
equi-distance filamental trajectory representation as mentioned in
Section \ref{sec:beyond-r0-sir}.{]}
\textcolor{violet}{We provide implementations of a few distance-based score functions that capture how "reasonable" an epidemic is relative to other epidemics, and these scores can be turned into an extremeness measure with}
\texttt{mean(sim\_scores\ \textgreater{}\ truth\_score)}.
{[}Specifically, functions like
\texttt{distance\_pseudo\_density\_function()} can calculate a
pseudo-density estimate of the true epidemic relative to simulated ones.
Functions \texttt{distance\_depth\_function()} and
\texttt{local\_distance\_depth\_function()} provide depth scores that
suggest how geometrically central an epidemic is to simulations.{]}
